---
title: "Surrey Climate Commission"
author: "Prado"
date: "2023-02-05"
output:
  github_document: default
  pdf_document: default
  html_document: default
---

Read the data from the UK Local authority
```{r}
datos<-read.csv("C:/Users/pradi/OneDrive/Desktop/ESG/uk-local-authority-ghg-emissions-2020-dataset.csv",header = T,sep=",")
head(datos)
str(datos)
```
```{r}
colnames(datos)<-c("Country","Country_Code","Region","Region_Code","Second_Tier_Authority",
                    "Local_Authority","Local_Authority_Code","Calendar_Year","LA_GHG_Sector",
                     "LA_GH_SubSector","Greenhouse_Gas","Territorial_Emissions_Kt_Co2",
                    "CO2_Emissions_Within_LA_Scope_KtCO2e",
                     "Mid_Year_Population_Thousands","Area_Km2")

```
###Analysis of Missing Data 

1) Filtering data for Surrey. I take the "Second Tier Authority" at Surrey Level data set and analysis missing data in the CO2 Emissions data. 

```{r}
library(dplyr)

surreyData<-datos%>%filter(Second_Tier_Authority=="Surrey")%>%select(Second_Tier_Authority,Local_Authority,Calendar_Year,LA_GHG_Sector,LA_GH_SubSector,Territorial_Emissions_Kt_Co2,CO2_Emissions_Within_LA_Scope_KtCO2e,Mid_Year_Population_Thousands,Area_Km2)

str(surreyData)
#df_status(surreyData)

```
There are no missing values or zeros so the database seems tidy.


2) aggregate total accumulated CO2 Emissions for each Second Tier Authority cross all the years
Give me Surrey's position ranked from top emiters to down

```{r}
library(ggplot2)
ordered<-datos%>%filter(Calendar_Year==2020, !(Second_Tier_Authority %in% c("Scotland","Wales","Northern Ireland")) )%>%group_by(Second_Tier_Authority)%>%summarise(Total=sum(Territorial_Emissions_Kt_Co2))%>%arrange(desc(Total))%>%mutate(ranked_LA=rank(-Total,))

Surrey_ranking<-subset(ordered[,c("Second_Tier_Authority","Total","ranked_LA")],Second_Tier_Authority=="Surrey")
Surrey_ranking

ggplot(head(ordered,20), aes(x=reorder(Second_Tier_Authority,-Total), y=Total,fill=Second_Tier_Authority))+geom_bar(stat="identity")+
  scale_x_discrete(breaks = levels(ordered$Second_Tier_Authority))+
       xlab("Second Authority Level") + 
       ylab ("Total CO2 emissions") 


```



Aggregate Emissions per Second tier Authority per Years 2018, 2019 and 2020

```{r}
emission_sector<-datos%>%filter(Calendar_Year %in% c(2018,2019,2020),Second_Tier_Authority=="Surrey")%>%
               group_by(Calendar_Year,Second_Tier_Authority,LA_GHG_Sector)%>%
               summarise(Total_CO2=sum(Territorial_Emissions_Kt_Co2))  
emission_sector

datos%>%filter(Calendar_Year %in% c(2018,2019,2020),Second_Tier_Authority=="Surrey")%>%
               group_by(Calendar_Year)%>%
               summarise(Total_CO2=sum(Territorial_Emissions_Kt_Co2)) 

```

Plot the difference in CO2 emitted per GHG Sector for the three years 2018, 2019,2020

```{r}
femission_sector<-emission_sector%>%filter(LA_GHG_Sector %in% c("Industry", "Commercial","Agriculture","Domestic","Transport","LULUCF"))%>% mutate(year=factor(Calendar_Year), FLA_GHG_Sector=factor(LA_GHG_Sector))

ggplot(femission_sector,aes(x=year,y=Total_CO2,group=FLA_GHG_Sector,color=FLA_GHG_Sector,alpha=1))+geom_line(size=2)+
    theme_classic()+
    geom_point( size = 2)+ geom_label(aes(label = round(Total_CO2,1)), 
             size = 3, 
             label.padding = unit(0.1, "lines"), 
             label.size = 0.0)

library(ggplot2)
ggplot(femission_sector,aes(x=FLA_GHG_Sector,y=Total_CO2,fill=year))+
   geom_bar(stat="identity", position = "dodge")+  
  geom_label(aes(label = round(Total_CO2,0)),colour = "black", size = 3,vjust = 0, position = position_dodge(1))+labs(title="Surrey Sector CO2 Emissions 2018-2020", x="Sector",y="Total CO2 Emissions",colour = "Cylinders", caption = "(based on data from  ONS Local Authority CO2 emissions estimates 2005-2020 (kt CO2) )")+theme_classic()+scale_fill_brewer()




```

# total CO2 aggregated by Year and energy type (LA_GHG_Subsector), 
```{r}
#Filtramos los datos para Surrey
Energy_Type<-surreyData%>%filter(Calendar_Year %in% c(2018,2019,2020))%>%group_by(Calendar_Year,LA_GHG_Sector,LA_GH_SubSector)%>%summarise(Total_energy_Type=round(sum(Territorial_Emissions_Kt_Co2),2))

#Create a new variable with the type of Energy consumed, Electricity, Gas, Fuel or Others

Energy_filtered<-Energy_Type%>%mutate(electricity_type=ifelse(grepl("Electricity",LA_GH_SubSector)==TRUE,"Electricity",
                                            ifelse(grepl("Gas",LA_GH_SubSector)==TRUE,"Gas", 
                                                   ifelse(LA_GH_SubSector %in% c("Road Transport (A roads)",  "Road Transport (Minor roads)","Road Transport (Motorways)"),"Fuel", ifelse(grepl("'Other'",LA_GH_SubSector)==TRUE,"Others_1","Others_2" )))))

Energy_filtered

#AGroup by Sector so we drop the subsectors.

Energy_aggregated<-Energy_filtered%>%group_by(Calendar_Year,LA_GHG_Sector,electricity_type)%>%summarise(Total=sum(Total_energy_Type))


#Finally, Agregamos por Year y por Tipo de electricidad. It coincides with the report!
energ_type_agg<-aggregate(Total~as.factor(Calendar_Year)+as.factor(electricity_type),data =Energy_aggregated,sum)




```


```{r}
#Using Separate function we can create a better variable as above but faster and more efficiently and accurate
library(tidytable)

enerFilt<-Energy_Type%>%separate(LA_GH_SubSector,into = c("GHG_Sector","Energy_Type"),sep=" ")
head(enerFilt)

EnerAgg<-enerFilt%>%group_by(Calendar_Year,LA_GHG_Sector,Energy_Type)%>%summarise(Total=sum(Total_energy_Type))
EnerAgg

Energy_agg<-EnerAgg%>%filter(Energy_Type %in% c("Electricity","Gas","'Other'","Transport"))%>%group_by(Energy_Type,Calendar_Year)%>%summarise(Total_CO2_EnergyType=sum(Total,na.rm=T))
Energy_agg

##Create a ggplot

ggplot(Energy_agg,aes(x=Energy_Type,y=Total_CO2_EnergyType,fill=as.factor(Calendar_Year)))+
   geom_bar(stat="identity", position = "dodge")+ labs(title="Surrey CO2 Emissions by Energy Type 2018-2020", x="Energy Type",y="Total CO2 Emissions",colour = "Cylinders", caption = "(based on data from  ONS Local Authority CO2 emissions estimates 2005-2020 (kt CO2)", fill="Year")+ 
  geom_label(aes(label = round(Total_CO2_EnergyType,0)),colour = "black", size = 3,vjust = 0, position = position_dodge(1))+theme_classic()+scale_fill_brewer()


```



#Surrey Emission by District and Borough area

```{r}
distEmissions<-surreyData%>%filter(Calendar_Year %in% c(2020))%>%group_by(Calendar_Year,Local_Authority)%>%summarise(Total_CO2=sum(Territorial_Emissions_Kt_Co2))

ggplot(distEmissions,aes(x=Total_CO2,y=reorder(Local_Authority,Total_CO2)))+
   geom_bar(stat="identity",color="blue", fill="blue")+  labs(title="Surrey CO2 Emissions by Local Authority 2020 (Total = 5494 KtCO2)", x="Total CO2 Emissions",y="Local Authority",colour = "Cylinders", caption = "(based on data from  ONS Local Authority CO2 emissions estimates 2005-2020 (kt CO2))", fill="Local Authority")+ 
  geom_label(aes(label = round(Total_CO2,0)),colour = "black", size = 3,vjust = 0)+theme_classic()

```
```{r}
distEmissionsTogether<-surreyData%>%filter(Calendar_Year %in% c(2018,2020))%>%group_by(Calendar_Year,Local_Authority)%>%summarise(Total_CO2=sum(Territorial_Emissions_Kt_Co2))


ggplot(distEmissionsTogether,aes(x=Total_CO2,y=reorder(Local_Authority,Total_CO2),fill=Calendar_Year))+
   geom_bar(stat="identity", position = position_dodge(width=.9))+ 
  geom_label(aes(label = round(Total_CO2,0)),colour = "white", size = 3,vjust = 0, position = position_dodge(1))+labs(title="Surrey LA CO2 Emissions 2020 vs. 2018", x="Total CO2",y="Local Authority",colour = "Year")+guides(fill = FALSE) +theme_classic()




```

